---
title: "Untitled"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output: html_document
---

# African Pollen Database


```{r conditional_print, child='rmarkdownSupportFiles/duplicateMaps.Rmd', eval = exists("map_dups_2")}


```

```{r rez1, echo=FALSE,message=FALSE,warning=FALSE,eval=exists("map_dups_2")}

if (exists("map_dups_2")) {
  cat(paste("A map of sites with duplicate geographies:"))
map_dups_2
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_2[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
datatable(dups_na_2,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_2")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_2
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_2[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
datatable(analymiss_2,rownames=FALSE)
} else {
      cat(paste("No missing analysis units"))
  }



```



# European Pollen Database

```{r rez_3, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_3")) {
  cat(paste("A map of sites with duplicate geographies:"))
map_dups_3
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_3[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
datatable(dups_na_3,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_3")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_3
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_3[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
datatable(analymiss_3,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }



```



# Indo-Pacific Pollen Database

```{r rez_4, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_4")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_4
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_4[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_4,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_4")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_4
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_4[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_4,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Latin American Pollen Database

```{r rez_5, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_5")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_5
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_5[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_5,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_5")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_5
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_5[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_5,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```


# North American Pollen Database

```{r rez_6, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_6")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_6
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_6[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_6,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_6")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_6
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_6[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_6,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```


# Pollen Database of Siberia and the Russian Far East


```{r rez_7, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_7")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_7
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_7[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_7,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_7")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_7
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_7[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_7,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# FAUNMAP 

```{r rez_10, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_10")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_10
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_10[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_10,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_10")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_10
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_10[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_10,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma (undifferentiated)

```{r rez_11, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_11")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_11
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_11[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_11,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_11")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_11
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_11[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_11,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# North American Plant Macrofossil Database

```{r rez_12, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_12")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_12
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_12[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_12,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_12")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_12
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_12[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_12,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Academy of Natural Sciences of Drexel University

```{r rez_13, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_13")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_13
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_13[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_13,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_13")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_13
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_13[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_13,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# NDSU Insect Database

```{r rez_14, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_14")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_14
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_14[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_14,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_14")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_14
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_14[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_14,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# North American Non-Marine Ostracode Database Project (NANODe)

```{r rez_15, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_15")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_15
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_15[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_15,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_15")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_15
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_15[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_15,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Alaskan Archaeofaunas

```{r rez_17, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_17")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_17
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_17[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_17,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_17")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_17
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_17[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_17,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# French Institute of Pondicherry Palynology and Paleoecology Database

```{r rez_18, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_18")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_18
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_18[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_18,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_18")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_18
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_18[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_18,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Japanese Pollen Database

```{r rez_19, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_19")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_19
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_19[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_19,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_19")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_19
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_19[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_19,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma Midden Database

```{r rez_20, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_20")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_20
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_20[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_20,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_20")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_20
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_20[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_20,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Chinese Pollen Database


```{r rez_22, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_22")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_22
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_22[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_22,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_22")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_22
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_22[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_22,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Holocene Perspective on Peatland Biogeochemistry

```{r rez_23, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_23")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_23
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_23[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_23,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_23")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_23
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_23[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_23,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma Testate Amoebae Database

```{r rez_25, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_25")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_25
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_25[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_25,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_25")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_25
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_25[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_25,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```



# Deep-Time Palynology Database

```{r rez_26, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_26")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_26
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_26[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_26,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_26")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_26
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_26[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_26,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma Biomarker Database

```{r rez_27, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_27")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_27
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_27[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_27,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_27")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_27
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_27[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_27,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Alpine Palynological Database

```{r rez_28, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_28")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_28
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_28[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_28,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_28")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_28
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_28[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_28,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Canadian Museum of Nature-Delorme Ostracoda- Surface Samples

```{r rez_29, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_29")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_29
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_29[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_29,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_29")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_29
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_29[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_29,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Diatom Paleolimnology Data Cooperative (DPDC)

```{r rez_30, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_30")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_30
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_30[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_30,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_30")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_30
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_30[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_30,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma Ostracode Database

```{r rez_31, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_31")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_31
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_31[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_31,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_31")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_31
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_31[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_31,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Faunal Isotope Database

```{r rez_32, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_32")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_32
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_32[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_32,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_32")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_32
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_32[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_32,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Neotoma Charcoal Data

```{r rez_33, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_33")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_33
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_33[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_33,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_33")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_33
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_33[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_33,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Pollen Monitoring Programme


```{r rez_35, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_35")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_35
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_35[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_35,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_35")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_35
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_35[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_35,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# PaleoVertebrates of Latin America


```{r rez_36, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_36")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_36
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_36[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_36,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_36")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_36
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_36[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_36,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# St. Croix Watershed Research Station of the Science Museum of Minnesota

```{r rez_37, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_37")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_37
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_37[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_37,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_37")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_37
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_37[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_37,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Tropical South American Diatom Database


```{r rez_38, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_38")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_38
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_38[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_38,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_38")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_38
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_38[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_38,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Marine Dinoflagellates Database

```{r rez_39, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_39")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_39
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_39[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_39,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_39")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_39
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_39[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_39,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Nonmarine Ostracod Distribution in Europe Database

```{r rez_41, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_41")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_41
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_41[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_41,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_41")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_41
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_41[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_41,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# East Asian Nonmarine Ostracod Database

```{r rez_42, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_42")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_42
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_42[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_42,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_42")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_42
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_42[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_42,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```

# Strategic Environmental Archaeology Database

```{r rez_43, echo=FALSE,message=FALSE,warning=FALSE}

if (exists("map_dups_43")) {
  cat(paste("A map of sites with duplicate geographies:"))
  map_dups_43
} else {
  cat(paste("No sites with duplicate geographies"))
}

if (length(dups_na_43[[1]]) > 0) {
   cat(paste("A table of sites with null geographies"))
   datatable(dups_na_43,rownames=FALSE)
} else {
    cat(paste("No sites with null geographies"))
  }

if (exists("whereanaly_43")) {
cat(paste("A map of sites missing analysis units:"))
  whereanaly_43
} else {
    cat(paste("No sites missing analysis units"))
  }


if (length(analymiss_43[[1]] > 0)) {
   cat(paste("A table of missing analysis units:"))
   datatable(analymiss_43,rownames=FALSE)
} else {
       cat(paste("No missing analysis units"))
  }


```
