---
title: "Bounding Box Sites in Neotoma"
author: "Nick Hoffman"
date: '2024-09-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Table of areas for bounding box sites (area in square Hectometers)

Histogram of areas in square meters, separated in to size classes to make easier to visualize.
```{r pressure, echo=FALSE,warning=FALSE,message=FALSE}
## https://hub.arcgis.com/datasets/esrica-tsg::indigenous-lands-of-canada/about

database_list = c("Cooperative Holocene Mapping Project", "African Pollen Database", "European Pollen Database", "Indo-Pacific Pollen Database", "Latin American Pollen Database", "North American Pollen Database", "Pollen Database of Siberia and the Russian Far East", "Canadian Pollen Database", "FAUNMAP", "Neotoma", "North American Plant Macrofossil Database", "Academy of Natural Sciences of Drexel University", "NDSU Insect Database", "North American Non-Marine Ostracode Database Project (NANODe)", "MioMap", "Alaskan Archaeofaunas", "French Institute of Pondicherry Palynology and Paleoecology Database", "Japanese Pollen Database", "Neotoma Midden Database", "Chinese Pollen Database", "Holocene Perspective on Peatland Biogeochemistry", "ANTIGUA", "Neotoma Testate Amoebae Database", "Deep-Time Palynology Database", "Neotoma Biomarker Database", "Alpine Pollen Database", "Canadian Museum of Nature-Delorme Ostracoda-Surface Samples", "Diatom Paleolimnology Data Cooperative (DPDC)", "Neotoma Ostracode Database", "Faunal Isotope Database", "Neotoma Charcoal Data", "Pollen Monitoring Programme", "PaleoVertebrates of Latin America", "St. Croix Watershed Research Station of the Science Museum of Minnesota","Marine Dinoflagellates Database","Tropical South American Diatom Database", "Packrat Middens", "Sedimentary aDNA Database")


library(DT)
library(tidyverse)
library(sf)
library(httr)
library(jsonlite)
library(geojsonsf)
library(units)
library(tmap)
library(rosm)
library(osmdata)
library(dplyr)


setwd("C:/Users/Nick/Documents")
rezes = read_sf("tl_2019_us_aiannh.shp") %>% dplyr::select(NAME)

canada_rezes = geojson_sf('https://proxyinternet.nrcan.gc.ca/arcgis/rest/services/CLSS-SATC/CLSS_Administrative_Boundaries/MapServer/0/query?outFields=*&where=1%3D1&f=geojson')
canada_rezes = canada_rezes %>% dplyr::rename('NAME' = 'adminAreaNameEng') %>% dplyr::select(NAME)
canada_rezes = st_transform(canada_rezes,crs="NAD83")
canada_rezes = canada_rezes %>% dplyr::mutate(NAME = str_to_title(NAME))

aust_ipa = geojson_sf("https://gis.environment.gov.au/gispubmap/rest/services/ogc_services/Indigenous_Protected_Areas/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")
aust_ipa = aust_ipa %>% dplyr::select(NAME)
aust_ipa = st_transform(aust_ipa,crs="NAD83")

rezes = rezes %>% rbind(canada_rezes) %>% rbind(aust_ipa)



idx=0
for (i in seq(1,length(database_list))) {
  #print(i)
  db_format = gsub(" ", '%20', database_list[i], fixed=TRUE)
  returns = content(GET(paste0("https://api.neotomadb.org/v2.0/data/datasets/db?limit=10000&offset=0&database=",db_format)))$data
  if(length(returns) != 0) {

    for (j in seq(length(returns))) {
      
      if(!is.null(returns[[j]]$site$geography)) {
              if (st_geometry_type(geojson_sf(returns[[j]]$site$geography)) != 'POINT') {
        for (k in seq(length(returns[[j]]$site$datasets))) {
          idx = idx + 1
        }
        }
    }
  }}
  
}


site_mat = matrix(nrow=idx,ncol=5)
idx=0
for (i in seq(1,length(database_list))) {
  db_format = gsub(" ", '%20', database_list[i], fixed=TRUE)
  returns = content(GET(paste0("https://api.neotomadb.org/v2.0/data/datasets/db?limit=10000&offset=0&database=",db_format)))$data
  if(length(returns) != 0) {
    for (j in seq(length(returns))) {
      if(!is.null(returns[[j]]$site$geography)) {
              if (st_geometry_type(geojson_sf(returns[[j]]$site$geography)) != 'POINT') {
        for (k in seq(length(returns[[j]]$site$datasets))) {
          idx = idx + 1
          if(!is.null(returns[[j]]$site$sitename )) {
            site_mat[idx,1] = returns[[j]]$site$sitename }
          if(!is.null(returns[[j]]$site$geography )) {
            site_mat[idx,2] = returns[[j]]$site$geography }
          if(!is.null(returns[[j]]$site$siteid )) {
            site_mat[idx,3] = returns[[j]]$site$siteid }
          if(!is.null(returns[[j]]$site$datasets[[k]]$datasettype )) {
            site_mat[idx,4] = returns[[j]]$site$datasets[[k]]$datasettype}
          if(!is.null(returns[[j]]$site$datasets[[k]]$database )) {
            site_mat[idx,5] = returns[[j]]$site$datasets[[k]]$database}
        }
        }
        
    }
  }
  }
}


site_df = as.data.frame(site_mat)
names(site_df)= c("site_name","geography","siteid","datasettype","database")

site_df1 = geojson_sf(site_df$geography) %>% cbind(site_df)

areas = site_df1  %>% distinct(site_name,siteid, geometry)


test = st_area(areas) %>% as.data.frame()
names(test) = c("area")

areas = areas %>% cbind(test)

library(ggplot2)

areas = areas %>% mutate(size = case_when(
  area< units::set_units(5*10^(5), m^2) ~ "small",
  area >= units::set_units(5*10^(5), m^2) & area< units::set_units(5*10^(7), m^2) ~ "medium",
  area >= units::set_units(5*10^(7), m^2) & area< units::set_units(5*10^(9), m^2) ~ "large",
  area >= units::set_units(5*10^(9), m^2) & area< units::set_units(5*10^(90), m^2) ~ "largest"
                         
))



areas = areas %>% mutate(size=fct_relevel(size,c("small","medium","large","largest"))) %>%
 arrange(size)


area_table = areas %>% dplyr::mutate(area = round(area/10000,2))

names(area_table) = c("sitename","siteid","area (hM^2)","size","geometry")

datatable(st_drop_geometry(area_table[1:4])) %>% formatCurrency("area (hM^2)",currency = "", interval = 3, mark = ",")

ggplot(areas) + 
  geom_histogram(mapping=aes(x=area),bins=600) +
  facet_wrap(~size,scales="free") +
  theme_bw() + 
  ggtitle("areas of bounding box sites in Neotoma")


#tmap_mode("view")
#tm_shape(osm.raster(areas)) + tm_rgb() +
#  tm_shape(test) + tm_fill(col=colors) +
#  tm_shape(areas) + tm_borders(lwd=2) 


  

library(leaflet) 

reps = ceiling(length(rezes[[1]])/11)
colors = rep(c("red","blue","limegreen","purple","yellow","black","orange","skyblue","pink","brown","darkgreen"),reps)

colors = colors[1:length(rezes[[1]])]
library(rmapshaper)
test = ms_simplify(rezes)
test = st_transform(test,crs="WGS84")


leaflet() %>%
     addTiles() %>%
   addPolygons(data = test,
        stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5,
        fillColor = ~colors,
        popup= ~NAME) %>%
    addPolygons(data = areas,
              stroke=TRUE,fillOpacity=0.1,fillColor = "white",smoothFactor=0.5,
              popup= ~site_name, color="black",weight="1") 


library(ggplot2)

dataset_areas = site_df1 %>% left_join(st_drop_geometry(areas))

fuzz_dataset_count= dataset_areas %>% group_by(datasettype) %>% count() %>% st_drop_geometry() %>% arrange(desc(n))

datatable(fuzz_dataset_count,rownames=FALSE)

over_ten = fuzz_dataset_count %>% dplyr::filter(n >= 10)


```


Below is a table of all the bounding box sites on (or intersecting) federal Indigenous lands.
```{r third, echo=FALSE,warning=FALSE}


rezes = st_transform(rezes,crs='WGS84')
sf_use_s2(FALSE)
sites_on_rez = st_join(areas,rezes) %>% drop_na(NAME)

datatable(st_drop_geometry(sites_on_rez))

```

Are sites that intersect reservations differently sized from sites that don't?

```{r fourth,echo=FALSE,warning=FALSE}

not_on_rez  = areas %>% dplyr::filter(!(siteid %in% sites_on_rez$siteid)) %>% mutate(class="not on federal Indigenous land")
on_rez = sites_on_rez %>% mutate(class="on federal Indigenous land") %>% dplyr::select(!NAME)

rez_comp = not_on_rez %>% rbind(on_rez)

ggplot(rez_comp) +
  geom_boxplot(mapping=aes(x=log(area),y=class)) + theme_bw()


t.test(as.vector(on_rez$area),as.vector(not_on_rez$area))
```

The graph below summarizes areas by datasettypes, only for those datasets for which there are more than ten bounding box sites.
```{r second, echo=FALSE,warning=FALSE}

meds = dataset_areas %>% group_by(datasettype) %>% summarize(med = median(area)) %>% arrange(desc(med))

dataset_areas = dataset_areas %>% dplyr::mutate(datasettype = fct_relevel(datasettype,st_drop_geometry(meds[[1]]))) %>%
 arrange(datasettype)

over_ten_dataset_areas = dataset_areas %>% dplyr::filter(datasettype %in% over_ten$datasettype)
ggplot(over_ten_dataset_areas) +
  geom_boxplot(mapping=aes(x=log(area),y=datasettype)) + theme_bw()


```



Dataset types on federal Indigenous lands versus others?

``` {r fifth, echo=FALSE, warning=FALSE}

dataset_rez_comp = left_join(over_ten_dataset_areas,st_drop_geometry(rez_comp))


ggplot(dataset_rez_comp) + 
  geom_bar(mapping=aes(x=class,fill=datasettype),position="fill")


chi_mat_prep = dataset_rez_comp %>% group_by(class, datasettype) %>% count() %>% st_drop_geometry() %>% pivot_wider(names_from="class",values_from="n")

datatable(chi_mat_prep,rownames=FALSE)

chi_mat_prep = chi_mat_prep %>% dplyr::mutate(`on federal Indigenous land` = case_when(
                                        is.na(`on federal Indigenous land`) ~ 0,
                                        TRUE ~ `on federal Indigenous land`
))
names(chi_mat_prep)= c("datasettype","notIndig","Indig")

chi_mat = chi_mat_prep[c(2,3)]

rownames(chi_mat) = chi_mat_prep$datasettype


library(graphics)
mosaicplot(chi_mat, shade = TRUE, las=2,
           main = "mosaic plot - dataset types by land tenure")


chisq = chisq.test(chi_mat)

library(corrplot)
corrplot(chisq$residuals, is.cor = FALSE)

areas_by_type_tenure = dataset_rez_comp %>% group_by(datasettype,class) %>% summarize(median = median(area)) %>% st_drop_geometry() %>% arrange(desc(median)) %>% pivot_wider(names_from="class",values_from="median")

areas_by_type_tenure = areas_by_type_tenure %>% mutate(`on federal Indigenous land` = as.vector(`on federal Indigenous land`)) %>% dplyr::mutate(`not on federal Indigenous land` = as.vector(`not on federal Indigenous land`))


areas_by_type_tenure= areas_by_type_tenure %>% dplyr::mutate(`on federal Indigenous land` = case_when(
                                        is.na(`on federal Indigenous land`) ~ 0,
                                        TRUE ~ `on federal Indigenous land`
))
names(areas_by_type_tenure)= c("datasettype","Indig","notIndig")

areas_by_type_tenure1 = areas_by_type_tenure[c(2,3)]

rownames(areas_by_type_tenure1) = areas_by_type_tenure$datasettype


mosaicplot(areas_by_type_tenure1, shade = TRUE, las=2,
           main = "mosaic plot - dataset types by land tenure, area not count")

chisq_area = chisq.test(areas_by_type_tenure1)


corrplot(chisq_area$residuals, is.cor = FALSE)


```