---
title: "Ireland Neotoma Metadata"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: no
    toc: true
    number_sections: true
    toc_depth: 1
    toc_float: true
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
h2, h3, h4, h5, h6 {
  counter-reset: section;
}
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

# Library Download

First I download all the libraries I'll need

```{r libraries}

library(tmap)
library(leaflet)
library(sf)
library(dplyr)
library(DT)
library(geojsonsf)
library(osmdata)
library(rosm)
library(httr)
```

# Sites

I create a bounding box around Ireland and make sure that it fits with a tmap plot.

```{r getsites_boundingbox}

lats = c(56, 56, 51, 51)
lons = c(-12, -5, -5, -12) 

# Create a data frame with coordinates
coordinates = data.frame(lat = lats, lon = lons)

# Convert to sf object and create a polygon
coordinates_sf = coordinates %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

# Plot to check
tm_shape(osm.raster(coordinates_sf)) +
  tm_rgb() +
  tm_shape(coordinates_sf) +
  tm_polygons(alpha = 0.5)
```

I feed the bounding box into one of the [Neotoma APIs](https://api.neotomadb.org/api-docs/) to return all the sites in that location. And I turn that into a dataframe.

```{r getsites_api}

coord_json = sf_geojson(coordinates_sf)

sites = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites?loc=",coord_json,"&limit=9999&offset=0")))$data


idx = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx = idx + 1
    }
    }
}


sites_mat = matrix(nrow=idx,ncol=11)

idx2 = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx2 = idx2 + 1
    for (m in seq(5)) {
      if (!is.null(sites[[i]][[m]])) {
        sites_mat[[idx2, m]] = sites[[i]][[m]]
      }
    }
    
     if (!is.null(sites[[i]]$collectionunits[[j]]$handle)) {
        sites_mat[[idx2,6]] = sites[[i]]$collectionunits[[j]]$handle
     }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunit)) {
        sites_mat[[idx2,7]] = sites[[i]]$collectionunits[[j]]$collectionunit
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunitid)) {
        sites_mat[[idx2,8]] = sites[[i]]$collectionunits[[j]]$collectionunitid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunittype)) {
        sites_mat[[idx2,9]] = sites[[i]]$collectionunits[[j]]$collectionunittype
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid)) {
        sites_mat[[idx2,10]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype)) {
        sites_mat[[idx2,11]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype
       }
    }
  }
}

sites_ir_df = as.data.frame(sites_mat)

names(sites_ir_df) = c("siteid","sitename","sitedescription","geography","altitude","handle","collectionunit","collectionunitid","collectionunittype","datasetid","datasettype")

sites_ir_sf = geojson_sf(sites_ir_df$geography) %>% cbind(sites_ir_df)

datasetids = sites_ir_df %>% dplyr::distinct(datasetid) 

#datasets_neo = get_datasets(as.numeric(datasetids$datasetid),all_data=TRUE)

#data = samples(get_downloads(datasets_neo,all_data=TRUE))

datatable(sites_ir_df,rownames=FALSE)

sites_ir_sf_unique = sites_ir_sf %>% distinct(siteid,.keep_all=TRUE)

 pointSites = sites_ir_sf_unique[st_geometry_type(sites_ir_sf_unique) == "POINT",] 
    polySites = sites_ir_sf_unique[st_geometry_type(sites_ir_sf_unique) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)
```

There are `r length(sites_ir_sf_unique[[1]])` sites in Ireland in Neotoma, associated with `r length(sites_ir_sf[[1]])` datasets.

# Old, Young Age
Next I use a different API to grab the oldest and youngest year associated with each dataset. These are probably in calibrated radiocarbon years...

```{r oldyoung}
idx=0
for (h in seq(length(datasetids[[1]]))) {
  returns = content(GET(paste0("https://api.neotomadb.org/v1.5/data/datasets/",datasetids$datasetid[[h]])))$data
for (i in seq(length(returns))) {
idx = idx +1
}}


oldyoung_mat = matrix(nrow=idx,ncol=3)

idx=0
for (h in seq(length(datasetids[[1]]))) {
  returns = content(GET(paste0("https://api.neotomadb.org/v1.5/data/datasets/",datasetids$datasetid[[h]])))$data
for (i in seq(length(returns))) {
  idx = idx + 1
  if (!is.null(returns[[i]]$ageoldest)) {
  oldyoung_mat[[idx,1]] = returns[[i]]$ageoldest}
  if (!is.null(returns[[i]]$ageyoungest)) {
   oldyoung_mat[[idx,2]] = returns[[i]]$ageyoungest}
  if (!is.null(returns[[i]]$datasetid)) {
   oldyoung_mat[[idx,3]]  = returns[[i]]$datasetid}
}}

oldyoung_df = as.data.frame(oldyoung_mat) 
names(oldyoung_df) = c("old","young","datasetid")

oldyoung_df = oldyoung_df %>% dplyr::mutate(datasetid = as.character(datasetid))
```

# Pubs

I download the database tables I'll need below. Datasetpublications links datasetids and publicationids, while publications has metadata for any given publication.

```{r downloadneotables}


tables = c("datasetpublications","publications")


for (i in seq(length(tables))) {

  table = tables[[i]]
list = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data


if (table != "publications") {
df = matrix(nrow=length(list),ncol=length(list[[1]]))

for (j in seq(1,length(list))) {
  for (k in seq(length(list[[1]]))) {
    if (!is.null(list[[j]][[k]])) {
      df[j,k] = list[[j]][[k]]
    }
  }
}


df = as.data.frame(df)

names(df) = names(list[[1]])  } else {
  df = matrix(nrow=length(list),ncol=(length(list[[1]])-2))

for (j in seq(1,length(list))) {
  for (k in seq((length(list[[1]])-2))) {
    if (!is.null(list[[j]][[k]])) {
      df[j,k] = list[[j]][[k]]
    }
  }
}
df = as.data.frame(df)

names(df) = names(list[[1]][1:(length(list[[1]])-2)]) }




assign(paste0(table,"_df"), df)
}

```

# Put it all together

I join all my different metadata together. Before I do that, I grab the x/y version of the coordinates instead of the sf version, which doesn't render well in tables.

```{r joins}

coords_polycenter = st_coordinates(st_centroid(polySites)) %>% cbind(st_drop_geometry(polySites))
coords_pointcenter = st_coordinates(pointSites) %>% cbind(st_drop_geometry(pointSites))

all_coords_df = rbind(coords_polycenter,coords_pointcenter)

joiner = sites_ir_df %>% left_join(datasetpublications_df, by=join_by(datasetid)) %>% left_join(publications_df, by=join_by(publicationid)) %>% left_join(oldyoung_df, by=join_by(datasetid)) %>% left_join(all_coords_df)

ask = joiner %>% select(c(publicationid, citation, sitename, X,Y, altitude, datasettype,old,young))

datatable(ask,rownames=FALSE)
```
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>