---
title: "Ireland Neotoma Metadata"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: no
    toc: true
    number_sections: true
    toc_depth: 1
    toc_float: true
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
h2, h3, h4, h5, h6 {
  counter-reset: section;
}
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

# Library Download

First I download all the libraries I'll need

```{r libraries}

library(tmap)
library(leaflet)
library(sf)
library(dplyr)
library(DT)
library(geojsonsf)
library(osmdata)
library(rosm)
library(httr)
```

# Sites

I create a bounding box around Ireland and make sure that it fits with a tmap plot.

```{r getsites_boundingbox}

lats = c(56, 56, 51, 51)
lons = c(-12, -5, -5, -12) 

# Create a data frame with coordinates
coordinates = data.frame(lat = lats, lon = lons)

# Convert to sf object and create a polygon
coordinates_sf = coordinates %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

# Plot to check
tm_shape(osm.raster(coordinates_sf)) +
  tm_rgb() +
  tm_shape(coordinates_sf) +
  tm_polygons(alpha = 0.5)
```

I feed the bounding box into one of the [Neotoma APIs](https://api.neotomadb.org/api-docs/) to return all the sites in that location. And I turn that into a dataframe.

```{r getsites_api}

coord_json = sf_geojson(coordinates_sf)

sites = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites?loc=",coord_json,"&limit=9999&offset=0")))$data


idx = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx = idx + 1
    }
    }
}


sites_mat = matrix(nrow=idx,ncol=11)

idx2 = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx2 = idx2 + 1
    for (m in seq(5)) {
      if (!is.null(sites[[i]][[m]])) {
        sites_mat[[idx2, m]] = sites[[i]][[m]]
      }
    }
    
     if (!is.null(sites[[i]]$collectionunits[[j]]$handle)) {
        sites_mat[[idx2,6]] = sites[[i]]$collectionunits[[j]]$handle
     }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunit)) {
        sites_mat[[idx2,7]] = sites[[i]]$collectionunits[[j]]$collectionunit
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunitid)) {
        sites_mat[[idx2,8]] = sites[[i]]$collectionunits[[j]]$collectionunitid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunittype)) {
        sites_mat[[idx2,9]] = sites[[i]]$collectionunits[[j]]$collectionunittype
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid)) {
        sites_mat[[idx2,10]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype)) {
        sites_mat[[idx2,11]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype
       }
    }
  }
}

sites_ir_df = as.data.frame(sites_mat)

names(sites_ir_df) = c("siteid","sitename","sitedescription","geography","altitude","handle","collectionunit","collectionunitid","collectionunittype","datasetid","datasettype")

sites_ir_sf = geojson_sf(sites_ir_df$geography) %>% cbind(sites_ir_df)

datasetids = sites_ir_df %>% dplyr::distinct(datasetid) 

#datasets_neo = get_datasets(as.numeric(datasetids$datasetid),all_data=TRUE)

#data = samples(get_downloads(datasets_neo,all_data=TRUE))

datatable(sites_ir_df,rownames=FALSE)

sites_ir_sf_unique = sites_ir_sf %>% distinct(siteid,.keep_all=TRUE)

 pointSites = sites_ir_sf_unique[st_geometry_type(sites_ir_sf_unique) == "POINT",] 
    polySites = sites_ir_sf_unique[st_geometry_type(sites_ir_sf_unique) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)
```

  
There are `r length(sites_ir_sf_unique[[1]])` sites in Ireland in Neotoma, associated with `r length(sites_ir_sf[[1]])` datasets. Below I count them by dataset type.

``` {r datasettypetable}

sites_ir_df %>% group_by(datasettype) %>% count() %>% arrange(desc(n)) %>% datatable(rownames=FALSE)
```

# Old, Young Age
Next I use a different API to grab the oldest and youngest year associated with each dataset. In Neotoma, a single site can be associated with multiple datasets, which have types like pollen, charcoal, cladocera...

```{r oldyoung}

site_string = paste0(distinct(sites_ir_sf_unique,siteid)$siteid, collapse=",")
returns = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites/",site_string,"/datasets?limit=2000")))$data

idx=0
for (i in seq(length(returns))) {
  for (j in seq(length(returns[[i]]$site$datasets))) {
    idx = idx +1
  }
}



oldyoung_mat = matrix(nrow=idx,ncol=4)

for (i in seq(length(returns))) {
  for (j in seq(length(returns[[i]]$site$datasets))) {
    if (!is.null(returns[[i]]$site$datasets[[j]]$agerange[[1]]$units)) {
      oldyoung_mat[[idx,1]] = returns[[i]]$site$datasets[[j]]$agerange[[1]]$units}
    if (!is.null(returns[[i]]$site$datasets[[j]]$agerange[[1]]$ageold)) {
      oldyoung_mat[[idx,2]] = returns[[i]]$site$datasets[[j]]$agerange[[1]]$ageold}
    if (!is.null(returns[[i]]$site$datasets[[j]]$agerange[[1]]$ageyoung)) {
      oldyoung_mat[[idx,3]]  = returns[[i]]$site$datasets[[j]]$agerange[[1]]$ageold}
    if (!is.null(returns[[i]]$site$datasets[[j]]$datasetid)) {
      oldyoung_mat[[idx,4]]  = returns[[i]]$site$datasets[[j]]$datasetid}
    
  }
}

oldyoung_df = as.data.frame(oldyoung_mat) 
names(oldyoung_df) = c("units","old","young","datasetid")

oldyoung_df = oldyoung_df %>% dplyr::mutate(datasetid = as.character(datasetid))
```

# Pubs

I download the publication tables. Datasetpublications links datasetids and publicationids, while publications has metadata for any given publication. In Neotoma, there is a many-to-many relationship between datasets and publications: many datasets can be associated with a particular publication, and vice versa.

```{r downloadneotables}


tables = c("datasetpublications","publications")


for (i in seq(length(tables))) {

  table = tables[[i]]
list = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data


if (table != "publications") {
df = matrix(nrow=length(list),ncol=length(list[[1]]))

for (j in seq(1,length(list))) {
  for (k in seq(length(list[[1]]))) {
    if (!is.null(list[[j]][[k]])) {
      df[j,k] = list[[j]][[k]]
    }
  }
}


df = as.data.frame(df)

names(df) = names(list[[1]])  } else {
  df = matrix(nrow=length(list),ncol=(length(list[[1]])-2))

for (j in seq(1,length(list))) {
  for (k in seq((length(list[[1]])-2))) {
    if (!is.null(list[[j]][[k]])) {
      df[j,k] = list[[j]][[k]]
    }
  }
}
df = as.data.frame(df)

names(df) = names(list[[1]][1:(length(list[[1]])-2)]) }




assign(paste0(table,"_df"), df)
}

```

# Put it all together

I join all my different metadata together. Before I do that, I grab the x/y version of the coordinates instead of the sf version, which doesn't render well in tables. Because of the one-to-many and many-to-many relationships among Neotoma tables, there are a lot of records in the below data table. But only about 100 Neotoma sites in Ireland, associated with about 300 datasets, and 64 publications. (And of those 300ish datasets, 62 are geochronologic type, which is a little artifactual, but I'm including it below for completeness.)

```{r joins}

coords_polycenter = st_coordinates(st_centroid(polySites)) %>% cbind(st_drop_geometry(polySites))
coords_pointcenter = st_coordinates(pointSites) %>% cbind(st_drop_geometry(pointSites))

all_coords_df = rbind(coords_polycenter,coords_pointcenter)

joiner = sites_ir_df %>% left_join(datasetpublications_df, by=join_by(datasetid)) %>% left_join(publications_df, by=join_by(publicationid)) %>% left_join(oldyoung_df, by=join_by(datasetid)) %>% left_join(all_coords_df)

ask = joiner %>% select(c(publicationid, citation, sitename, X,Y, altitude, datasettype,old,young, units))

datatable(ask,rownames=FALSE)
```
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>