---
title: "River Basin Survey sites in Neotoma"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: no
    toc: true
    number_sections: true
    toc_depth: 1
    toc_float: true
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
h2, h3, h4, h5, h6 {
  counter-reset: section;
}
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

```{r setup, warning=FALSE,message=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(dplyr)
library(DT)
library(geojsonsf)
library(osmdata)
library(rosm)
library(rmapshaper)
library(tmap)
library(sf)
library(leaflet)
library(stringr)



states = read_sf("tl_2024_us_state.shp")
states = ms_simplify(states)

# https://www.digtech-llc.com/blog/2012/3/27/94-shovelbums-guide-part-15-the-smithsonian-trinomial-system.html

state_names = c("Alabama", "Alaska", "American Samoa", "Arizona", "Arkansas", "California", "Colorado","Connecticut","Delaware", "District of Columbia", "Federated States of Micronesia", "Florida", "Georgia","Guam","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Marshall Islands", "Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Northern Mariana Islands","Ohio","Oklahoma","Oregon","Palau","Pennsylvania","Puerto Rico","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virgin Islands","Virginia","Washington","West Virginia","Wisconsin","Wyoming")

tris = c(1,49,NA,2,3,4,5,6,7,NA,NA,8,9,NA,50,10,11,12,13,14,15,16,17,NA,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,NA,33,34,35,NA,36,NA,37,38,39,40,41,42,43,NA,44,45,46,47,48)


state_tri_mapper=data.frame(names=state_names,trinom_num=tris) %>% dplyr::mutate(trinom_num = as.character(trinom_num))

state_tri_mapper = state_tri_mapper %>% dplyr::mutate(trinom_num = case_when(names=="Arizona" ~ "AZ", names=="California" ~ "CA", names=="New Mexico" ~ "LA",names=="Maine" ~ "ME",
         TRUE ~ trinom_num))


siteid=c("junk")
sitename=c("junk")
state=c("junk")
collectionunitid=c("junk")
geography=("junk")
potentials = data.frame(siteid=siteid,sitename=sitename,state=state,geography=geography, collectionunitid=collectionunitid)
for (num in seq(length(states[[1]]))) {

coordinates_sf = states[num,] %>% st_convex_hull()


coord_json = sf_geojson(coordinates_sf) %>% URLencode()


sites = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites?loc=",coord_json,"&limit=9999&offset=0")))$data

if (length(sites) > 0) {

idx = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx = idx + 1
    }
    }
}


sites_mat = matrix(nrow=idx,ncol=11)

idx2 = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx2 = idx2 + 1
    for (m in seq(5)) {
      if (!is.null(sites[[i]][[m]])) {
        sites_mat[[idx2, m]] = sites[[i]][[m]]
      }
    }
    
     if (!is.null(sites[[i]]$collectionunits[[j]]$handle)) {
        sites_mat[[idx2,6]] = sites[[i]]$collectionunits[[j]]$handle
     }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunit)) {
        sites_mat[[idx2,7]] = sites[[i]]$collectionunits[[j]]$collectionunit
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunitid)) {
        sites_mat[[idx2,8]] = sites[[i]]$collectionunits[[j]]$collectionunitid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunittype)) {
        sites_mat[[idx2,9]] = sites[[i]]$collectionunits[[j]]$collectionunittype
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid)) {
        sites_mat[[idx2,10]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype)) {
        sites_mat[[idx2,11]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype
       }
    }
  }
}

sites_state_df = as.data.frame(sites_mat)

names(sites_state_df) = c("siteid","sitename","sitedescription","geography","altitude","handle","collectionunit","collectionunitid","collectionunittype","datasetid","datasettype")

numb = state_tri_mapper %>% dplyr::filter(names == states$NAME[[num]]) %>% dplyr::select(trinom_num)

if (states$NAME[[num]] == "Vermont") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "VT")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (states$NAME[[num]] == "Rhode Island") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "RI")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (states$NAME[[num]] == "Connectictut") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "CT")| str_detect(sitename, "6")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (states$NAME[[num]] == "Arizona") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "[A-Za-z]{1,2}:[0-9]{1,2}")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (states$NAME[[num]] == "Alaska") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "[A-Za-z]{3}-[0-9]+")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (states$NAME[[num]] == "New York") {
  potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename, "\\d{5}")|
str_detect(sitename,"30")) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
  potentials = rbind(potentials,potentialsites)
} else if (!is.na(numb[[1]])) {
potentialsites = sites_state_df %>% dplyr::filter(str_detect(sitename,as.character(numb[[1]]))) %>% dplyr::select(siteid,sitename,geography,collectionunitid) %>% dplyr::mutate(state = states$NAME[[num]])
potentials = rbind(potentials,potentialsites)
  } else {
  #print(paste0("Null mapper value for ", states$NAME[[num]]))
}}}


potentials = potentials %>% dplyr::filter(sitename != "junk") %>% dplyr::filter(!str_detect(sitename,"Jensen 1998"))  %>% dplyr::filter(!str_detect(sitename,"Riegel 1965")) %>% dplyr::filter(!str_detect(sitename,"Haile 8A")) %>% dplyr::filter(!str_detect(sitename,"Santa Fe River 8A")) %>% distinct() %>% dplyr::filter(!sitename %in% c("Site 66 (Delcourt et al. 1983)","Site 110 (G.M. Peterson 1978)","Site 10 (Mack, Bryant, and Pell 1978)","LRRB31","Site 106 (Swain unpublished)","Anthony Cave [MALB 29]","Baldy Peak Cave [MALB 29]","Solar Energy Generating Station II [1.76.42]","Luz Solar Trough [1.76.34]", "Site 13 (Delcourt et al. 1983)","Site 14 (Delcourt et al. 1983)","Site 66 (Delcourt et al. 1983)","CHUSKA29","CHUS29","Upper Alturas [UO Loc 2424; CAS Loc 36805]","Site 35 (Heusser 1978)","BECL360","Stand 45 (Mack and Bryant 1974]","Law's [MSv100]","Seven Springs [CE101x3]","Rock House [MS201]","Mobile Bay, MB0810-GC20","Spring Canyon [A-41]","Pratt Cave [TMM-41172]","Little Sunday Canyon Local Fauna [V5414]","Avery Ranch [s:5:8]","Big Juniper House [1595]","Badger House [1453]","Mesa Verde Site 875","La Poudre LP5 Site","SNMN5","SNMN15","ANIM5","ANIM35","COMO5","LPLT5","GATC26","Mile 49 Lac la Ronge Highway","Stations 49-55 (Lichti-Federovich and Ritchie 1968)","TK-49","49 [HFL49]","Valkenswaard N49","Mosquito Lake Site, [PSM-00049]","BOREA24, Mile 49","LR49","ALTAK49","ABD249","Island 35 [10-Q-7]","Hancock Ditch (ditch #39)","Tunica Bayou Site 22","NLA06608-1232","JHMN24 (McAndrews and Wright 1969)","NLA06608-0439","Twin Peaks, AZ","Tucson Mountains, Pima Co., AZ","Vijdt Polder (B49D0349)","Borteldonk (B49F0399)","Halsteren (B49B0566)","Ossendrecht 49D20-1","Shangri-La Bog","Oskino-09","'White-2 (Le4) Pond'","Mud Pond (VT)")) %>%
  dplyr::filter(!str_detect(sitename,"Mount Evans Surface")) %>%  dplyr::filter(!str_detect(sitename,"Elliot-Fisk")) %>%  dplyr::filter(!str_detect(sitename,"Raux-Feui")) %>% dplyr::filter(!str_detect(sitename,"Exposure")) %>%
  dplyr::filter(!str_detect(sitename,"LDDsite")) %>%  dplyr::filter(!str_detect(sitename,"Pond,")) %>%
  dplyr::filter(!str_detect(sitename,"Begin boardwalk")) %>% dplyr::filter(!str_detect(sitename,"unpublished")) %>%
  dplyr::filter(!str_detect(sitename,"Colonel's Island Site")) %>%
  dplyr::filter(!str_detect(sitename,"Delcourt")) %>% dplyr::filter((state !="California") | state == "California" & str_detect(sitename,"CA"))


us_bounds = read_sf("cb_2018_us_nation_20m.shp")


unique_pot = geojson_sf(potentials$geography) %>% cbind(potentials) %>% distinct(siteid,.keep_all=TRUE) %>% st_transform(crs=st_crs(us_bounds)) %>% st_filter(us_bounds)
```

# Sites in Neotoma with Smithsonian Trinomials
```{r show, echo=FALSE,warning=FALSE,message=FALSE}
datatable(st_drop_geometry(unique_pot)[c('siteid','sitename','state')],rownames=FALSE)


 pointSites = unique_pot[st_geometry_type(unique_pot) == "POINT",] 
    polySites = unique_pot[st_geometry_type(unique_pot) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.5, 
                   stroke = FALSE, 
                   popup = ~sitename)

  

tables = c("datasets","publications", "datasetpublications","collectionunits")



for (i in seq(length(tables))) {

  table = tables[[i]]
dslinks = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data



if (table == "publications") {
  dsl_df = matrix(nrow=length(dslinks),ncol=(length(dslinks[[1]]) - 2))
  for (j in seq(1,length(dslinks))) {
  for (k in seq((length(dslinks[[1]])) - 2)) {
    if (!is.null(dslinks[[j]][[k]])) {
      dsl_df[j,k] = dslinks[[j]][[k]]
    }
  }
}
}
else {
  dsl_df = matrix(nrow=length(dslinks),ncol=length(dslinks[[1]]))
for (j in seq(1,length(dslinks))) {
  for (k in seq(length(dslinks[[1]]))) {
    if (!is.null(dslinks[[j]][[k]])) {
      dsl_df[j,k] = dslinks[[j]][[k]]
    }
  }
}
}

dsl_df = as.data.frame(dsl_df)

if (table =="publications") {
names(dsl_df) = names(dslinks[[1]])[1:(length(dslinks[[1]])-2)]  }

else {
  names(dsl_df) = names(dslinks[[1]])  
}

assign(paste0(table,"_df"), dsl_df)
}




```
# Publication Metadata

## Sites associated with papers from before 1967

The River Basin Surveys Program began in 1946 and ended in 1967. The below sites are all associated in Neotoma with at least one publication from 1967 or earlier, so these ones are most likely to have been connected to the program. Some of these (n=23) actually are published before the RBS even begins. But I include them here anyway.

``` {r publics, message=FALSE,warning=FALSE,echo=FALSE}


pubs = potentials %>% left_join(datasets_df, by=join_by(collectionunitid)) %>% left_join(datasetpublications_df, by=join_by(datasetid)) %>% left_join(publications_df, by=join_by(publicationid)) 


pre1967 = pubs %>% dplyr::filter(!is.na(citation)) %>% dplyr::filter(as.numeric(year)<=1967) %>% group_by(siteid) %>% summarize(siteid,sitename,state.x,toString(unique(year)),toString(unique(citation), sep="\n\n"),geography) %>% distinct()

pre1967_sf = geojson_sf(pre1967$geography) %>% cbind(pre1967) 


names(pre1967) = c("siteid","sitename","state","publication years","citations","geography")



pubyears = pubs %>% distinct(citation,.keep_all = TRUE) %>% select(year)

pre1967 %>% select(!geography) %>% datatable(rownames=FALSE)



pointSites_1967 = pre1967_sf[st_geometry_type(pre1967_sf) == "POINT",] 
polySites_1967 = pre1967_sf[st_geometry_type(pre1967_sf) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites_1967, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites_1967, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.5, 
                   stroke = FALSE, 
                   popup = ~sitename)

  

```
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>