---
title: "South Dakota Neotoma"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: no
    toc: true
    number_sections: true
    toc_depth: 1
    toc_float: true
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
h2, h3, h4, h5, h6 {
  counter-reset: section;
}
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---
# SD wide
```{r setup, warning=FALSE,message=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(dplyr)
library(DT)
library(geojsonsf)
library(osmdata)
library(rosm)
library(rmapshaper)
library(tmap)
library(sf)
library(leaflet)
library(lubridate)
library(neotoma2)

#lats = c(46, 46, 42.42, 42.42)
#lons = c(-104.3, -96, -96, -104.3) 
lats <- c(45.944106, 45.933153, 45.818137, 45.604536, 45.412843, 45.297827, 43.501391, 43.479483, 43.397329, 43.222067, 43.123482, 43.052282, 42.855112, 42.707235, 42.488157, 42.515542, 42.657942, 42.844158, 42.844158, 42.866066, 42.767481, 42.94822, 42.997512, 42.997512, 43.002989, 43.002989, 44.996596, 44.996596, 45.944106)

lons <- c(-104.047534, -96.560556, -96.582464, -96.856311, -96.681049, -96.451017, -96.451017, -96.582464, -96.527695, -96.560556, -96.434587, -96.511264, -96.544125, -96.631756, -96.44554, -96.626279, -96.692003, -97.217789, -97.688806, -97.831206, -97.951699, -98.466531, -98.499393, -101.626726, -103.324578, -104.053011, -104.058488, -104.042057, -104.047534)

# Create a data frame with coordinates
coordinates = data.frame(lat = lats, lon = lons)

# Convert to sf object and create a polygon
coordinates_sf = coordinates %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

# Plot to check
tm_shape(osm.raster(coordinates_sf)) +
  tm_rgb() +
  tm_shape(coordinates_sf) +
  tm_polygons(alpha = 0.5)



coord_json = sf_geojson(coordinates_sf)




sites = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites?loc=",coord_json,"&limit=9999&offset=0")))$data


idx = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx = idx + 1
    }
    }
}


sites_mat = matrix(nrow=idx,ncol=11)

idx2 = 0
for (i in seq(length(sites))) {
  for (j in seq(length(sites[[i]]$collectionunits))) {
    for (k in seq(length(sites[[i]]$collectionunits[[j]]$datasets))) {
    idx2 = idx2 + 1
    for (m in seq(5)) {
      if (!is.null(sites[[i]][[m]])) {
        sites_mat[[idx2, m]] = sites[[i]][[m]]
      }
    }
    
     if (!is.null(sites[[i]]$collectionunits[[j]]$handle)) {
        sites_mat[[idx2,6]] = sites[[i]]$collectionunits[[j]]$handle
     }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunit)) {
        sites_mat[[idx2,7]] = sites[[i]]$collectionunits[[j]]$collectionunit
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunitid)) {
        sites_mat[[idx2,8]] = sites[[i]]$collectionunits[[j]]$collectionunitid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$collectionunittype)) {
        sites_mat[[idx2,9]] = sites[[i]]$collectionunits[[j]]$collectionunittype
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid)) {
        sites_mat[[idx2,10]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasetid
       }
       if (!is.null(sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype)) {
        sites_mat[[idx2,11]] = sites[[i]]$collectionunits[[j]]$dataset[[k]]$datasettype
       }
    }
  }
}

sites_sd_df = as.data.frame(sites_mat)

names(sites_sd_df) = c("siteid","sitename","sitedescription","geography","altitude","handle","collectionunit","collectionunitid","collectionunittype","datasetid","datasettype")

sites_sd_sf = geojson_sf(sites_sd_df$geography) %>% cbind(sites_sd_df)

datasetids = sites_sd_df %>% dplyr::distinct(datasetid) 

#datasets_neo = get_datasets(as.numeric(datasetids$datasetid),all_data=TRUE)

#data = samples(get_downloads(datasets_neo,all_data=TRUE))

datatable(sites_sd_df,rownames=FALSE)

sites_sd_sf_unique = sites_sd_sf %>% distinct(siteid,.keep_all=TRUE)

 pointSites = sites_sd_sf_unique[st_geometry_type(sites_sd_sf_unique) == "POINT",] 
    polySites = sites_sd_sf_unique[st_geometry_type(sites_sd_sf_unique) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)

  
datatable(st_drop_geometry(sites_sd_sf_unique),rownames=FALSE)

rezes = read_sf("tl_2019_us_aiannh.shp")

coordinates_sf = st_transform(coordinates_sf,crs=st_crs(rezes))

sd_rezes = rezes %>% dplyr::select(NAME) %>% st_filter(coordinates_sf)


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = sd_rezes, 
              color = "black", 
              weight = 5, 
              fillColor = "white", 
              fillOpacity = 0.35, 
              popup = ~NAME) %>%
  addPolygons(data = polySites, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)
  

sites_sd_sf_unique = st_transform(sites_sd_sf_unique,crs=st_crs(rezes))

tribal_sites = sites_sd_sf_unique %>% st_filter(sd_rezes)
```
# Tribal Sites
```{r chunl2, message=FALSE,echo=FALSE,warning=FALSE}

datatable(st_drop_geometry(tribal_sites),rownames=FALSE)
```

# RBS sites
```{r rbs, message=FALSE,warning=FALSE,echo=FALSE}

trinomials = sites_sd_sf_unique$sitename[grep("39",sites_sd_sf_unique$sitename)]


smiths = sites_sd_sf_unique %>% dplyr::filter(sitename %in% trinomials) %>% dplyr::filter(sitename!= "NLA06608-0439")


 pointSites_s = smiths[st_geometry_type(smiths) == "POINT",] 
    polySites_s = smiths[st_geometry_type(smiths) == "POLYGON",] 


  leaflet() %>%
  addTiles() %>%
  addPolygons(data = sd_rezes, 
              color = "black", 
              weight = 5, 
              fillColor = "white", 
              fillOpacity = 0.35, 
              popup = ~NAME) %>%
  addPolygons(data = polySites, 
              color = "blue", 
              weight = 5, 
              fillColor = "blue", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename) %>%
  addPolygons(data = polySites_s, 
              color = "red", 
              weight = 5, 
              fillColor = "red", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites_s, 
                   radius = 5, 
                   color = "red", 
                   fillColor = "red", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)
  
  
smiths_tribal = smiths %>% st_filter(sd_rezes) 

datatable(st_drop_geometry(smiths),rownames=FALSE)

#E,warning=FALSE,message=FALSE}


tables = c("publications", "datasetpublications","sites","dataprocessors","collectionunits","datasets","contacts")



for (i in seq(length(tables))) {

  table = tables[[i]]
dslinks = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data



if (table == "publications") {
  dsl_df = matrix(nrow=length(dslinks),ncol=(length(dslinks[[1]]) - 2))
  for (j in seq(1,length(dslinks))) {
  for (k in seq((length(dslinks[[1]])) - 2)) {
    if (!is.null(dslinks[[j]][[k]])) {
      dsl_df[j,k] = dslinks[[j]][[k]]
    }
  }
}
}
else {
  dsl_df = matrix(nrow=length(dslinks),ncol=length(dslinks[[1]]))
for (j in seq(1,length(dslinks))) {
  for (k in seq(length(dslinks[[1]]))) {
    if (!is.null(dslinks[[j]][[k]])) {
      dsl_df[j,k] = dslinks[[j]][[k]]
    }
  }
}
}

dsl_df = as.data.frame(dsl_df)

if (table =="publications") {
names(dsl_df) = names(dslinks[[1]])[1:(length(dslinks[[1]])-2)]  }

else {
  names(dsl_df) = names(dslinks[[1]])  
}

assign(paste0(table,"_df"), dsl_df)
}

smithpubs = datasetpublications_df %>% dplyr::filter(datasetid %in% smiths$datasetid) %>% left_join(publications_df, by=join_by(publicationid))

datatable(smithpubs,rownames=FALSE)

```

# RBS Sites on federally recognized reservations lands in SD
```{r lasty, message=FALSE,warning=FALSE,echo=FALSE}
smiths_tribal %>% st_drop_geometry() %>% datatable(rownames=FALSE)

```
# Sites added Later

Two sites were added in 2018. One of these (Blaine) is not an RBS site. The other (Medicine Creek) is an RBS site. Medicine Creek was added by Jordan Schutz in July of that year. It's not recorded who added Blaine in April.

```{r sitemata, echo=FALSE,message=FALSE,warning=FALSE}


latesites1 = sites_df %>% dplyr::filter(siteid %in% smiths$siteid) %>% mutate(date = as.Date(recdatecreated)) %>% dplyr::filter(year(date) >2013)

latesites2 = smiths %>% dplyr::filter(siteid %in% latesites1$siteid) %>% left_join(collectionunits_df,by=join_by(collectionunitid))  %>% left_join(datasets_df,by=join_by(datasetid))

dataprocessors_df %>% dplyr::filter(datasetid %in% latesites2$datasetid) %>% dplyr::left_join(contacts_df,by=join_by(contactid))

datatable(latesites1,rownames=FALSE)

```

# Number species

## Number of specimens for Semken Falk

All but two Semken Falk sites have greater than 11 species

``` {r checker, echo=FALSE,message=FALSE,warning=FALSE}

rbsmap = read.csv("rbs_semken_sd.csv") %>% dplyr::select(1,2,4,5) %>% dplyr::mutate(siteid=as.character(siteid))

smith_datasets =smiths %>% left_join(collectionunits_df,by=join_by(collectionunitid))  %>% left_join(datasets_df,by=join_by(datasetid))

neodata = get_datasets(c(as.numeric(smith_datasets$datasetid)),all_data=TRUE)

samps = neodata %>% get_downloads(all_data=TRUE) %>% samples()

distinct_species = samps %>% group_by(siteid) %>% distinct(variablename) %>% ungroup() %>% count(siteid) %>% left_join(rbsmap)

sem = distinct_species %>% dplyr::filter(Semken..Falk.== "yes") %>% arrange(desc(n))

datatable(sem,rownames=FALSE)
```

## Non Semken Falk

Only 6 of the non Semken Falk sites have greater than 11 species!

```{r nonsem, echo=FALSE,message=FALSE,warning=FALSE}
nosem = distinct_species %>% dplyr::filter(Semken..Falk.== "no") %>% arrange(desc(n))

datatable(nosem,rownames=FALSE)

```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>