---
title: "Neotoma's Human Records"
author: "Nick Hoffman"
date: "October 31st 2023"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: yes
    number_sections: no
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

This document shows what records of Homo sapiens I was able to find in Neotoma. My method was first to download the entire "taxa" table (46,302 rows) and find the taxonid for any relevant sounding taxa. The first table below shows the taxa I found.


```{r setup, echo=FALSE,include=TRUE,message = FALSE}

library(neotoma2)
library(DT)
library(sf)
library(tidyverse)
library(httr)
library(jsonlite)
library(tmap)
library(osmdata)
library(rosm)
library(geojsonsf)
library(stringr)

## get taxa
taxalist = content(GET("https://api.neotomadb.org/v2.0/data/dbtables?table=taxa&limit=75000&offset=0"))$data

taxa_df = matrix(nrow=length(taxalist),ncol=14)

for (i in seq(1,length(taxalist))) {
  for (j in seq(1,14)) {
    if (!is.null(taxalist[[i]][[j]])) {
      taxa_df[i,j] = taxalist[[i]][[j]]
    }
  }
}

taxa_df = as.data.frame(taxa_df)

names(taxa_df) = names(taxalist[[1]])

test1 = grep("Homo",taxa_df$taxonname)

homos = taxa_df[c(5571,6261,6262,6631),]

datatable(homos[c(1,3,6,7)],rownames = FALSE)

## relevant indices are 5571, 6261, 6631

##higher taxa indices are 6261, 6262, 5623


## get occurrences from taxon id
homolist = content(GET("https://api.neotomadb.org/v2.0/data/taxa/6116,6821,6822,7196/occurrences?limit=2500&offset=0"))$data

```

(FYI, taxonid 6359 is Primates, and 6171 is Mammalia)

Then I searched for all occurrences of any of those taxa. 

The map below shows the sites they come from, and the table below that shows the sample information on those 28 human samples I found.

```{r occurrences, echo=FALSE,include=TRUE,message = FALSE}

homo_df = matrix(nrow=length(homolist),ncol=15)

for (i in seq(1,length(homolist))) {
  for (j in seq(1,1)) {
    if (!is.null(homolist[[i]][[j]])) {
      homo_df[i,j] = homolist[[i]][[j]]
    }
    
  }
   for (k in seq(1,4)) {
    if (!is.null(homolist[[i]][[2]][[k]])) {
      homo_df[i,(1+k)] = homolist[[i]][[2]][[k]]
    }}
   for (k in seq(1,3)) {
    if (!is.null(homolist[[i]][[3]][[k]])) {
      homo_df[i,(5+k)] = homolist[[i]][[3]][[k]]
    } }
  
   for (k in seq(1,7)) {
    if (!is.null(homolist[[i]][[4]][[k]])) {
      homo_df[i,(8+k)] = homolist[[i]][[4]][[k]]
    }}
}

homo_df = as.data.frame(homo_df)

names(homo_df) = c("occid","taxonid","taxonname","value","sampleunits","age","ageolder","ageyounger","datasetid","siteid","sitename","altitude","location","datasettype","database")

homo_sites = homo_df %>% dplyr::select(siteid) %>% distinct()

homo_sites_neo = get_sites(c(as.numeric(homo_sites[[1]])))

homo_sites_sf = as.data.frame(homo_sites_neo) %>%
  st_as_sf(coords=c("long","lat"),crs="+proj=longlat +datum=WGS84")


homo_bg = osm.raster(homo_sites_sf)


tm_shape(homo_bg)+
    tm_rgb() +
    tm_shape(homo_sites_sf) +
    tm_dots(size=0.1,alpha=0.4) +
    tm_layout(
              main.title= 'Sites with Homo Samples in Neotoma', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)


datatable(homo_df[c(1,3,4,5,7,8,9,10,11,15)],rownames=FALSE)
```

I think there must be different conventions going on here for the age ranges. All the negative age values haven't happened yet, if they're meant to correspond to "number of years before 1950." 


Below I just separate out the sites that have polygon-shapes when I searched for them with the occurrence API, out of curiosity...

```{r polygons, echo=FALSE,include=TRUE,message = FALSE}

poly = homo_df$occid[str_detect(homo_df$location,"Polygon")]

poly_df = homo_df %>% dplyr::filter(occid %in% poly)

poly_sf = rbind(geojson_sf(poly_df$location[1]),geojson_sf(poly_df$location[2]),geojson_sf(poly_df$location[3]),geojson_sf(poly_df$location[4]),geojson_sf(poly_df$location[5]))


poly_bg = osm.raster(poly_sf[[1]][2:4])


tm_shape(poly_bg)+
    tm_rgb() +
    tm_shape(poly_sf[[1]][2:3]) +
    tm_polygons(alpha=0.5,col="blue",fill="blue") +
    tm_layout(
              main.title= 'first three polygons', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)



long1 = as.data.frame(c(-79,-77))
lat1 = as.data.frame(c(35,38)) 

names(long1) = c("long1")
names(lat1) = c("lat1")
lat1 = lat1 %>%
  cbind(long1) %>%
  st_as_sf(coords=c("long1","lat1"),crs="+proj=longlat +datum=WGS84") %>%
  rbind(poly_sf[[1]][4])


poly_bg2 = osm.raster(lat1)

tm_shape(poly_bg2)+
    tm_rgb() +
    tm_shape(poly_sf[[1]][4]) +
    tm_polygons(alpha=0.5,col="blue",fill="blue") +
    tm_layout(
              main.title= 'fourth polygon', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)



long1 = as.data.frame(c(24,26))
lat1 = as.data.frame(c(39,41)) 

names(long1) = c("long1")
names(lat1) = c("lat1")
lat1 = lat1 %>%
  cbind(long1) %>%
  st_as_sf(coords=c("long1","lat1"),crs="+proj=longlat +datum=WGS84") %>%
  rbind(poly_sf[[1]][5])


poly_bg3 = osm.raster(lat1)

tm_shape(poly_bg3)+
    tm_rgb() +
    tm_shape(poly_sf[[1]][5]) +
    tm_polygons(alpha=0.5,col="blue",fill="blue") +
    tm_layout(
              main.title= 'fifth polygon', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)




```