---
title: "Neotoma's Human Records"
author: "Nick Hoffman"
date: "Updated: 3/12/2024"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: yes
    number_sections: no
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

This document shows what records of Homo sapiens I was able to find in Neotoma. My method was first to download the whole taxa table and find the taxon id for any relevant sounding taxa. The first table below shows the taxa I found.


```{r setup, echo=FALSE,include=TRUE,message = FALSE, warning=FALSE}

library(neotoma2)
library(DT)
library(sf)
library(tidyverse)
library(httr)
library(jsonlite)
library(tmap)
library(osmdata)
library(rosm)
library(geojsonsf)
library(stringr)

## get taxa
taxalist = content(GET("https://api.neotomadb.org/v2.0/data/dbtables?table=taxa&limit=75000&offset=0"))$data

taxa_df = matrix(nrow=length(taxalist),ncol=14)

for (i in seq(1,length(taxalist))) {
  for (j in seq(1,14)) {
    if (!is.null(taxalist[[i]][[j]])) {
      taxa_df[i,j] = taxalist[[i]][[j]]
    }
  }
}

taxa_df = as.data.frame(taxa_df)

names(taxa_df) = names(taxalist[[1]])

test1 = grep("Homo ",taxa_df$taxonname)

homos = taxa_df[c(grep("Homo ",taxa_df$taxonname),grep("^Homo$", taxa_df$taxonname),grep("^Homini", taxa_df$taxonname)),]

datatable(homos[c(1,3,6,7)],rownames = FALSE)

## relevant indices are 5571, 6261, 6631

##higher taxa indices are 6261, 6262, 5623


## get occurrences from taxon id
homolist = content(GET("https://api.neotomadb.org/v2.0/data/taxa/6116,6821,6822,7196/occurrences?limit=2500&offset=0"))$data

```

(FYI, taxonid 6359 is Primates, and 6171 is Mammalia)

Then I searched for all occurrences of any of those taxa. 

The two maps below (one static, one dynamic) shows the sites they come from, and the table below that shows the sample information on those 28 human samples I found at 22 distinct sites.

```{r occurrences, echo=FALSE,include=TRUE,message = FALSE}

homo_df = matrix(nrow=length(homolist),ncol=15)

for (i in seq(1,length(homolist))) {
  for (j in seq(1,1)) {
    if (!is.null(homolist[[i]][[j]])) {
      homo_df[i,j] = homolist[[i]][[j]]
    }
    
  }
   for (k in seq(1,4)) {
    if (!is.null(homolist[[i]][[2]][[k]])) {
      homo_df[i,(1+k)] = homolist[[i]][[2]][[k]]
    }}
   for (k in seq(1,3)) {
    if (!is.null(homolist[[i]][[3]][[k]])) {
      homo_df[i,(5+k)] = homolist[[i]][[3]][[k]]
    } }
  
   for (k in seq(1,7)) {
    if (!is.null(homolist[[i]][[4]][[k]])) {
      homo_df[i,(8+k)] = homolist[[i]][[4]][[k]]
    }}
}

homo_df = as.data.frame(homo_df)

names(homo_df) = c("occid","taxonid","taxonname","value","sampleunits","age","ageolder","ageyounger","datasetid","siteid","sitename","altitude","location","datasettype","database")

homo_sites = homo_df %>% dplyr::select(siteid) %>% distinct()

homo_sites_neo = get_sites(c(as.numeric(homo_sites[[1]])))

homo_sites_sf = as.data.frame(homo_sites_neo) %>%
  st_as_sf(coords=c("long","lat"),crs="+proj=longlat +datum=WGS84")

homo_full = left_join(homo_sites_sf,homo_df)

homo_bg = osm.raster(homo_sites_sf)


tm_shape(homo_bg)+
    tm_rgb() +
    tm_shape(homo_full) +
    tm_dots("database",size=0.1,alpha=0.4,palette=c('black','red','green',"purple")) +
    tm_layout(legend.position=c("RIGHT","BOTTOM"),
              legend.bg.color="white",
              legend.bg.alpha=0.9,
              legend.width=0.4,
              legend.text.size=0.5,
              main.title= 'Sites with Homo Samples in Neotoma', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)
```
<h3 style="text-align:center;">Human Ancestor Sites</h3>
``` {r another, echo=FALSE,include=TRUE,message = FALSE}
plotLeaflet(homo_sites_neo) 

datatable(homo_df[c(1,3,4,5,7,8,9,10,11,15)],rownames=FALSE)
```



Below I count the occurrences by database and taxonname.

```{r polygons, echo=FALSE,include=TRUE,message = FALSE}
counttabl = homo_df %>% group_by(database,taxonname) %>% count() %>% arrange(desc(n))

datatable(counttabl,rownames = FALSE)
```

Below I list all geopolitical units associated with any of the 22 distinct sites.


```{r countries, echo=FALSE,include=TRUE,message = FALSE}

sitestring = paste0(homo_df$siteid,collapse=",")

gps = content(GET(paste0("https://api.neotomadb.org/v2.0/data/sites/",sitestring,"/geopoliticalunits?limit=250&offset=0")))$data

gps_mat = matrix(nrow=52,ncol=6)
idx_sites = 0
for (i in seq(length(gps))) {
  for (j in seq(length(gps[[i]]$geopoliticalunits))) {
    idx_sites = idx_sites + 1
    if (!is.null(gps[[i]]$siteid)) {
    gps_mat[idx_sites,1] = gps[[i]]$siteid}
    if (!is.null(gps[[i]]$geopoliticalunits[[j]][[1]])) {
    gps_mat[idx_sites,2] = gps[[i]]$geopoliticalunits[[j]][[1]] }
    if (!is.null(gps[[i]]$geopoliticalunits[[j]][[2]])) {
    gps_mat[idx_sites,3] = gps[[i]]$geopoliticalunits[[j]][[2]] }
    if (!is.null(gps[[i]]$geopoliticalunits[[j]][[3]])) {
    gps_mat[idx_sites,4] = gps[[i]]$geopoliticalunits[[j]][[3]] }
    if (!is.null(gps[[i]]$geopoliticalunits[[j]][[4]])) {
    gps_mat[idx_sites,5] = gps[[i]]$geopoliticalunits[[j]][[4]] }
    if (!is.null(gps[[i]]$geopoliticalunits[[j]][[5]])) {
    gps_mat[idx_sites,6] = gps[[i]]$geopoliticalunits[[j]][[5]] }
  }
}


gps_df = as.data.frame(gps_mat)

names(gps_df) = c("siteid","rank","gp_id","gp_name","gp_unit","higher_gp_id")

datatable(gps_df,rownames=FALSE)

gps_sum = gps_df %>% dplyr::filter(rank==1) %>% group_by(gp_name) %>% count() %>% arrange(desc(n))
```

Lastly, I count the countries the distinct sites are in.

```{r highest, echo=FALSE,include=TRUE,message = FALSE}

datatable(gps_sum,rownames=FALSE)
```