---
title: "Neotoma's Panther Records"
author: "Nick Hoffman"
date: "November 3rd 2023"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: yes
    number_sections: no
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

This document shows what records of pumas I was able to find in Neotoma. My method was first to download the entire "taxa" table (46,302 rows) and find the taxonid for any relevant sounding taxa. The first table below shows the taxa I found.


```{r setup, echo=FALSE,include=TRUE,message = FALSE}

library(neotoma2)
library(DT)
library(sf)
library(tidyverse)
library(httr)
library(jsonlite)
library(tmap)
library(osmdata)
library(rosm)
library(geojsonsf)
library(stringr)

## get taxa
taxalist = content(GET("https://api.neotomadb.org/v2.0/data/dbtables?table=taxa&limit=75000&offset=0"))$data

taxa_df = matrix(nrow=length(taxalist),ncol=14)

for (i in seq(1,length(taxalist))) {
  for (j in seq(1,14)) {
    if (!is.null(taxalist[[i]][[j]])) {
      taxa_df[i,j] = taxalist[[i]][[j]]
    }
  }
}

taxa_df = as.data.frame(taxa_df)

names(taxa_df) = names(taxalist[[1]])

pumas = taxa_df[c(6186,6187,6188,6198,6616,7130,7131,9481,9485,13226,14530,15248),]

datatable(pumas[c(1,3,6,7)],rownames = FALSE)



## get occurrences from taxon id
pumalist = content(GET("https://api.neotomadb.org/v2.0/data/taxa/15248,6740,6741,6742,6751,7181,7692,7693,10091,10095,13861,15481/occurrences?limit=2500&offset=0"))$data

```

(FYI, taxonid 6063 is Felidae and taxonid 19271 is Carnivora incertae sedis)

Then I searched for all occurrences of any of those taxa. 

The two maps below (one static, one dynamic) shows the 151 sites they come from. 

The table below that shows the sample information on the 241 puma samples I found.

```{r occurrences, echo=FALSE,include=TRUE,message = FALSE}

puma_df = matrix(nrow=length(pumalist),ncol=15)

for (i in seq(1,length(pumalist))) {
  for (j in seq(1,1)) {
    if (!is.null(pumalist[[i]][[j]])) {
      puma_df[i,j] = pumalist[[i]][[j]]
    }
    
  }
   for (k in seq(1,4)) {
    if (!is.null(pumalist[[i]][[2]][[k]])) {
      puma_df[i,(1+k)] = pumalist[[i]][[2]][[k]]
    }}
   for (k in seq(1,3)) {
    if (!is.null(pumalist[[i]][[3]][[k]])) {
      puma_df[i,(5+k)] = pumalist[[i]][[3]][[k]]
    } }
  
   for (k in seq(1,7)) {
    if (!is.null(pumalist[[i]][[4]][[k]])) {
      puma_df[i,(8+k)] = pumalist[[i]][[4]][[k]]
    }}
}

puma_df = as.data.frame(puma_df)

names(puma_df) = c("occid","taxonid","taxonname","value","sampleunits","age","ageolder","ageyounger","datasetid","siteid","sitename","altitude","location","datasettype","database")

puma_sites = puma_df %>% dplyr::select(siteid) %>% distinct()

puma_sites_neo = get_sites(c(as.numeric(puma_sites[[1]])),all_data=TRUE)

puma_sites_sf = as.data.frame(puma_sites_neo) %>%
  st_as_sf(coords=c("long","lat"),crs="+proj=longlat +datum=WGS84")


lat1 = as.data.frame(c(-5))
long1 = as.data.frame(c(-50))
names(lat1) = c("lat1")
names(long1) = c("long1")
lat1 = lat1 %>%
  cbind(long1) %>%
  st_as_sf(coords=c("long1","lat1"),crs="+proj=longlat +datum=WGS84") %>%
  rbind(puma_sites_sf$geometry)


puma_bg = osm.raster(lat1)


join = left_join(puma_sites_sf,puma_df)

tm_shape(puma_bg)+
    tm_rgb() +
    tm_shape(join) +
    tm_dots("taxonname",size=0.1,alpha=0.4,palette="Set1") +
    tm_layout(legend.position=c("LEFT","BOTTOM"),
              legend.bg.color="white",
              legend.bg.alpha=0.9,
              legend.width=0.4,
              legend.text.size=0.9,
              main.title= 'Sites with Puma Samples in Neotoma', 
              main.title.position = "center",
              title.bg.color = "white", panel.label.height=1)
```
<h3 style="text-align:center;">Puma Sites</h3>
``` {r another, echo=FALSE,include=TRUE,message = FALSE}
plotLeaflet(puma_sites_neo) 

datatable(puma_df[c(1,3,4,5,7,8,9,10,11,15)],rownames=FALSE)
```

Below I count the occurrences by database and taxonname.

```{r polygons, echo=FALSE,include=TRUE,message = FALSE}
counttabl = puma_df %>% group_by(database,taxonname) %>% count() %>% arrange(desc(n))

datatable(counttabl,rownames = FALSE)
