---
title: "State of the Database by Constituent db"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output:
  html_document:
    css: db_by_db.css
    df_print: paged
    highlight: pygment
    keep_md: no
    theme: journal
editor_options:
    chunk_output_type: inline
---

```{r setup, include=TRUE, echo=FALSE, warning=FALSE,message=FALSE}
library(leaflet)
library(wkb)
library(dplyr)
library(sf)
library(httr)
library(DT)
library(htmltools)
library(tidyverse)
```

```{r, include=FALSE}
datatable(iris)
```

```{r download_analysisunits, message=FALSE,warning=FALSE,echo=FALSE}

length_analysisunits = content(GET("https://api.neotomadb.org/v2.0/data/dbtables/analysisunits?count=true"))$data
if (is.null(length_analysisunits)) {
  length_analysisunits = content(GET("https://api.neotomadb.org/v2.0/data/dbtables/analysisunits?count=true"))$data
}

num_reps = ceiling(as.numeric(length_analysisunits[[1]]$count)/80000)

analysisunits = list()
for (i in seq(num_reps)) {
  start= (i-1)*80000
  end = i*80000
  newnits = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/analysisunits?count=false&offset=",start,"&limit=",end)))$data
  
  analysisunits = append(analysisunits,newnits)
}


analysis_mat = matrix(nrow=length(analysisunits),ncol=11)

for (i in seq(length(analysisunits))) {
  for(j in seq(11) ) {
    if (!is.null(analysisunits[[i]][[j]])) {
      analysis_mat[i,j] = analysisunits[[i]][[j]]
    }}}


analysis_df = as.data.frame(analysis_mat)


names(analysis_df) = c("analysisunitid","collectionunitid","analysisunitname","depth","thickness","faciesid","mixed","igsn","notes","datecreated","datemodified")

```


```{r download-tables, echo=FALSE,message=FALSE,warning=FALSE}

tables = c("dslinks","sites", "datasetdatabases","constituentdatabases","collectionunits","datasets")


for (i in seq(length(tables))) {
  
  table = tables[[i]]
  dslinks = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data
  if (is.null(dslinks)) {
    #print("uh oh")
    #paste("uh oh")
    dslinks = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables?table=",table,"&limit=75000&offset=0")))$data
  }
  
  dsl_df = matrix(nrow=length(dslinks),ncol=length(dslinks[[1]]))
  
  for (j in seq(1,length(dslinks))) {
    for (k in seq(length(dslinks[[1]]))) {
      if (!is.null(dslinks[[j]][[k]])) {
        dsl_df[j,k] = dslinks[[j]][[k]]
      }
    }
  }
  
  dsl_df = as.data.frame(dsl_df)
  
  names(dsl_df) = names(dslinks[[1]]) 
  
  assign(paste0(table,"_df"), dsl_df)
}

sites_df = sites_df %>% dplyr::mutate(siteid = as.numeric(siteid))

collectionunits_df = collectionunits_df %>% dplyr::mutate(siteid = as.numeric(siteid))
dslinks_df = dslinks_df %>% dplyr::mutate(datasetid=as.character(datasetid))

```


``` {r firstquery, message=FALSE,warning=FALSE,echo=FALSE}

source('no_db_analysis.R')

source("figure_looper.R")
```
# Data anomalies by constituent database

This page provides information on anomalies in Neotoma, on a constituent database by database basis. You can click on the tab for any of the databases below (or 'no database' at the bottom) to receive information about duplicate site geographies, null site geographies, and sites without any connection to analysis units.

# {.tabset .tabset-pills .tabcontent}
```{r name, results="asis", echo=FALSE}
for (i in seq(length(db_list))) {
  cat("\n")
  cat(paste0("## ", db_list[[i]]$db_name, "\n"))
#  cat("<h1>", (i+1), db_list[[i]]$db_name, "</h1>\n")

cat("### Duplicate Geographies \n") 
if (!is.null(db_list[[i]]$dup_map)) {
   cat("\n")
  cat("<i>Note that in the duplicate geographies table, all locations are reported as points, though they may actually be stored as polygons in the database.</i>")
  cat("\n")
print(tagList(db_list[[i]]$dup_map))
  cat("\n")
print(tagList(datatable(db_list[[i]]$dups)))
} else {
  cat(paste("No sites with duplicate geographies \n"))
  cat("\n")
}

   cat("### Null Geographies \n")
if (length(db_list[[i]]$nulls[[1]] > 0)) {
   cat("\n")

print(tagList(datatable(db_list[[i]]$nulls, rownames=FALSE)))
  cat("\n")
} else {
    cat(paste("No sites with null geographies \n"))
    cat("\n")
  }
   cat("### Missing Analysis Units \n")
if (length(db_list[[i]]$missing_analysis[[1]]) > 0) {
cat(paste("A map of sites missing analysis units:"))
    cat("\n")
 print(tagList(db_list[[i]]$map_missing_analysis))
 print(tagList(datatable(db_list[[i]]$missing_analysis, rownames=FALSE)))
   cat("\n")
 
} else {
    cat(paste("No sites missing analysis units \n"))
    cat("\n")
  }

}
```


## No database

```{r newnie, message=FALSE, warning=FALSE,echo=FALSE}


joiner = dslinks_df %>% left_join(sites_df, by = join_by(siteid)) %>% left_join(datasetdatabases_df, by = join_by(datasetid)) %>% left_join(constituentdatabases_df) 

dups = joiner %>% group_by(siteid) %>% summarize(toString(unique(datasetid))) %>% left_join(sites_df) %>% group_by(geog) %>% count() %>% dplyr::filter(n>1) %>% left_join(joiner)

missers = sites_df %>% left_join(collectionunits_df,by=join_by(siteid)) %>% left_join(datasets_df,by=join_by(collectionunitid))  %>% left_join(datasetdatabases_df, by = join_by(datasetid)) %>% left_join(constituentdatabases_df) %>% dplyr::filter(is.na(collectionunitid) | is.na(datasetid)) %>% dplyr::select(c(siteid,sitename,collectionunitid,datasetid,databasename,databaseid,geog))


m_sfc = st_as_sfc(hex2raw(missers$geog),EWKB=TRUE)
missers_sf = st_as_sf(missers,geom=m_sfc) %>% dplyr::select(!geog)


missers_sf = missers_sf %>% dplyr::mutate(whatmiss = case_when(
  is.na(datasetid) & is.na(collectionunitid) ~ paste0(sitename, " missing: dataset + collectionunit"),
  is.na(datasetid) & !is.na(collectionunitid) ~ paste0(sitename, " missing: dataset"),
  !is.na(datasetid) & is.na(collectionunitid) ~ paste0(sitename, " missing: collectionunit"),
  !is.na(datasetid) & !is.na(collectionunitid) ~ paste0("none missing from ",sitename)))
                                                               

 pointSites_m = missers_sf[st_geometry_type(missers_sf) == "POINT",] 
    polySites_m = missers_sf[st_geometry_type(missers_sf) == "POLYGON",] 

  
analy_missers = sites_df %>% left_join(collectionunits_df,by=join_by(siteid)) %>% left_join(analysis_df, by=join_by(collectionunitid)) %>% left_join(datasets_df,by=join_by(collectionunitid))  %>% left_join(datasetdatabases_df, by = join_by(datasetid)) %>% left_join(constituentdatabases_df) %>% dplyr::filter(is.na(analysisunitid)) %>% dplyr::select(c(siteid,sitename,collectionunitid,datasetid,databasename,databaseid,analysisunitid, geog)) 


am_sfc = st_as_sfc(hex2raw(analy_missers$geog),EWKB=TRUE)
analy_missers_sf = st_as_sf(analy_missers,geom=am_sfc) %>% dplyr::select(!geog)


analy_missers_nodb = analy_missers_sf %>% dplyr::filter(is.na(databaseid))

print("A map of sites missing datasets and/or collection units:")
    leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites_m, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~whatmiss)  %>%
  addCircleMarkers(data = pointSites_m, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~whatmiss)

    

print("A table of sites missing datasets AND analysis units:")
datatable(analy_missers_nodb,rownames=FALSE)


print("A map of sites missing datasets AND analysis units:")                                             

 pointSites_am = analy_missers_nodb[st_geometry_type(analy_missers_nodb) == "POINT",] 
    polySites_am = analy_missers_nodb[st_geometry_type(analy_missers_nodb) == "POLYGON",] 

    leaflet() %>%
  addTiles() %>%
  addPolygons(data = polySites_am, 
              color = "red", 
              weight = 5, 
              fillColor = "orange", 
              fillOpacity = 0.35, 
              popup = ~sitename)  %>%
  addCircleMarkers(data = pointSites_am, 
                   radius = 5, 
                   color = "blue", 
                   fillColor = "blue", 
                   fillOpacity = 0.2, 
                   stroke = FALSE, 
                   popup = ~sitename)

```
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
