---
title: "Answers from Const DB Survey"
author: "Nick Hoffman"
date: "`r Sys.Date()`"
output: 
  html_document:
  df_print: paged
css: db_by_db.css
toc: true
number_sections: true
toc_depth: 1
toc_float: true
theme: journal
---
  
```{r libs, message=FALSE,warning=FALSE,echo=FALSE}
library(dplyr)
library(DT)

surv= read.csv("constdbsurvey.csv")

names(surv) = c("time","name","db","other_db","africa","arctic","farEast","China","Japan","otherAsia","atlantic","europe","centralAmerica","Alaska","NorthAmerica","Canada","Oceania","Pacific","SouthAmerica","geoscope_other","post1850","Quaternary","preQuaternary","timescope_other","biomarker","charcoal","diatom","dinoflag","insect","aDNA","ostracode","plantmacro","pollen","specstableisotope","testateamoebae","vertfauna","waterchem","biochem","chironomid","cladocera","geochem","loi","macroinvert","organiccarbon","paleomag","physsed","phytolith","stableisotope","xrayfluor","proxyscope_other","leadstewards_dbdescrip","dbdescrip","otherlinks_dbdescrip","newdata","wholead","qualdescribe_scope","arcticocean","atlanticocean","indianocean","pacificocean","southernocean","oceania","asia","antarctica","otherproxy_share","othertime_share","othergeo_share","changename")

surv = surv %>% dplyr::select(!c(arctic,farEast,China,Japan,otherAsia,atlantic,Alaska,Canada,Oceania,Pacific))
```


# Who answered?

```{r answers, echo=FALSE,warning=FALSE,message=FALSE}

surv=  surv %>% dplyr::mutate(db = case_when(
  other_db != "" & db == "Other" ~ other_db,
  other_db != "" & db != "Other" ~ paste0(db,", ", other_db),
  TRUE ~ db
)) %>% dplyr::select(!other_db)
surv %>% group_by(db) %>% summarize(name,wholead,qualdescribe_scope,changename,leadstewards_dbdescrip,dbdescrip,otherlinks_dbdescrip,newdata) %>% datatable(rownames=FALSE)
```

# Who didn't answer?

As of March 6, these have not answered:

* Pollen Database of Siberia and the Russian Far East
* FAUNMAP
* Neotoma
* NDSU Insect Database
* Japanese Pollen Database
* Neotoma Midden Database
* Chinese Pollen Database
* Holocene Perspective on Peatland Biogeochemistry
* Deep-Time Palynology Database
* Neotoma Biomarker Database
* Faunal Isotope Database
* Neotoma Charcoal Data (but PaleoFire did - are these the same or different?)
* Pollen Monitoring Programme
* Nonmarine Ostracod Distribution in Europe Database

# Geographic Scope

```{r geoscope, message=FALSE,warning=FALSE,echo=FALSE}
surv %>% dplyr::select(name,db,africa,europe,centralAmerica,NorthAmerica,SouthAmerica,oceania,asia,antarctica,arcticocean,atlanticocean,indianocean,pacificocean,southernocean,geoscope_other,othergeo_share) %>% datatable(rownames=FALSE)
```

## Temporal Scope

```{r tempscope, message=FALSE,warning=FALSE,echo=FALSE}

surv %>% dplyr::select(name,db,post1850,Quaternary,preQuaternary,timescope_other,othertime_share) %>% datatable(rownames=FALSE)
```

## Proxy Scope

```{r proxyscope, message=FALSE,warning=FALSE,echo=FALSE}

surv %>% dplyr::select(name,db,biomarker,charcoal,diatom,dinoflag,insect,aDNA,ostracode,plantmacro,pollen,specstableisotope,testateamoebae,vertfauna,waterchem,biochem,chironomid,cladocera,geochem,loi,macroinvert,organiccarbon,paleomag,physsed,phytolith,stableisotope,xrayfluor,proxyscope_other,otherproxy_share) %>% datatable(rownames=FALSE)
```